---
title: "House_price_EDA_doug"
author: "Doug Devens"
date: "5/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## read in file

```{r}
train_data = read.csv('train.csv', stringsAsFactors = FALSE)
str(train_data)
```
```{r}
summary(train_data)
```

## look for NAs
Not much missing data except for: ~40% fireplaces, ~2.5% basements, 94% alleys, 80% fences, ~2% lot frontage, ~5% garages and 99.5% pools.
```{r}
a = colSums(is.na.data.frame(train_data))
a[a>0]
```
Most houses in Ames wouldn't have pools, and most probably wouldn't have alleys except the older areas.  It's a bit of a surprise more houses don't have fences, but suburbs out here wouldn't have fences I think.  There was a period of time when houses didn't have fireplaces, but a lot of older houses and newer houses should have them.  Most houses have garages, but maybe not the older ones.
looking now at combinations to see where they're happening
```{r}
library(mice)
mice::md.pattern(train_data, rotate.names = TRUE)
```

## Relationships between input variables
1) see that missing lot frontage is highest in cul de sac lots, but not exclusively
```{r}
table(train_data$LotConfig, is.na(train_data$LotFrontage))
```
Also see HouseStyle and GrLivArea (above grade living area) seem to correlate though maybe it's only enough to say 1 story and two story are different based on square footage it seems
```{r}
library(ggplot2)
ggplot(train_data, aes(x=HouseStyle, y=GrLivArea)) + geom_boxplot()
```

House style and whether there is a fireplace appear to correlate
```{r}
table(train_data$HouseStyle, is.na(train_data$FireplaceQu))
```
begin looking at correlations
```{r}
library(dplyr)
```
time series of yr sold
```{r}
train_data %>% group_by(., YrSold) %>% summarise(., mean(SalePrice))
```
 ## start changing column types and count number of NAs
```{r}
train_data$MSSubClass = as.character(train_data$MSSubClass)
train_data$LotShape = ifelse(train_data$LotShape=='Reg',1,
                      ifelse(train_data$LotShape=='IR1',2,
                      ifelse(train_data$LotShape=='IR2',3,4)))
a = colSums(is.na.data.frame(train_data))
a[a>0]
```
continue changing what I think are ordinal strings to numbers so we can judge correlation against other parameters
```{r}
train_data$Utilities = ifelse(train_data$Utilities=='AllPub',1,
                      ifelse(train_data$Utilities=='NoSewr',2,
                      ifelse(train_data$Utilities=='NoSeWa',3,4)))
train_data$LandSlope = ifelse(train_data$LandSlope=='Gtl',1,
                      ifelse(train_data$LandSlope=='Mod',2,3))
train_data$ExterQual = ifelse(train_data$ExterQual=='Ex',1,
                      ifelse(train_data$ExterQual=='Gd',2,
                      ifelse(train_data$ExterQual=='TA',3,
                      ifelse(train_data$ExterQual=='Fa',4,5))))
train_data$ExterCond = ifelse(train_data$ExterCond=='Ex',1,
                      ifelse(train_data$ExterCond=='Gd',2,
                      ifelse(train_data$ExterCond=='TA',3,
                      ifelse(train_data$ExterCond=='Fa',4,5))))
train_data$BsmtQual = ifelse(is.na(train_data$BsmtQual),6,
                      ifelse(train_data$BsmtQual=='Ex',1,
                      ifelse(train_data$BsmtQual=='Gd',2,
                      ifelse(train_data$BsmtQual=='TA',3,
                      ifelse(train_data$BsmtQual=='Fa',4,5)))))
train_data$BsmtCond = ifelse(is.na(train_data$BsmtCond),6,
                      ifelse(train_data$BsmtCond=='Ex',1,
                      ifelse(train_data$BsmtCond=='Gd',2,
                      ifelse(train_data$BsmtCond=='TA',3,
                      ifelse(train_data$BsmtCond=='Fa',4,5)))))
train_data$BsmtExposure = ifelse(is.na(train_data$BsmtExposure),5,
                      ifelse(train_data$BsmtExposure=='Gd',1,
                      ifelse(train_data$BsmtExposure=='Av',2,
                      ifelse(train_data$BsmtExposure=='Mn',3,4))))
train_data$BsmtFinType1 = ifelse(is.na(train_data$BsmtFinType1),7,
                      ifelse(train_data$BsmtFinType1=='GLQ',1,
                      ifelse(train_data$BsmtFinType2=='ALQ',2,
                      ifelse(train_data$BsmtFinType1=='BLQ',3,
                      ifelse(train_data$BsmtFinType1=='Rec',4,
                      ifelse(train_data$BsmtFinType1=='LwQ',5,6))))))
train_data$HeatingQC = ifelse(train_data$HeatingQC=='Ex',1,
                      ifelse(train_data$HeatingQC=='Gd',2,
                      ifelse(train_data$HeatingQC=='TA',3,
                      ifelse(train_data$HeatingQC=='Fa',4,5))))
train_data$Electrical = ifelse(is.na(train_data$Electrical),3,
                      ifelse(train_data$Electrical=='SBrkr',1,
                      ifelse(train_data$Electrical=='FuseA',2,
                      ifelse(train_data$Electrical=='FuseF',3,
                      ifelse(train_data$Electrical=='FuseP',4,3)))))
train_data$KitchenQual = ifelse(train_data$KitchenQual=='Ex',1,
                      ifelse(train_data$KitchenQual=='Gd',2,
                      ifelse(train_data$KitchenQual=='TA',3,
                      ifelse(train_data$KitchenQual=='Fa',4,5))))
train_data$Functional = ifelse(train_data$Functional=='Typ',1,
                      ifelse(train_data$Functional=='Min1',2,
                      ifelse(train_data$Functional=='Min2',3,
                      ifelse(train_data$Functional=='Mod',4,
                      ifelse(train_data$Functional=='Maj1',5,
                      ifelse(train_data$Functional=='Maj2',6,
                      ifelse(train_data$Functional=='Sev',7,8)))))))
train_data$FireplaceQu = ifelse(is.na(train_data$FireplaceQu),6,
                      ifelse(train_data$FireplaceQu=='Ex',1,
                      ifelse(train_data$FireplaceQu=='Gd',2,
                      ifelse(train_data$FireplaceQu=='TA',3,
                      ifelse(train_data$FireplaceQu=='Fa',4,5)))))
train_data$GarageFinish = ifelse(is.na(train_data$GarageFinish),4,
                      ifelse(train_data$GarageFinish=='Fin',1,
                      ifelse(train_data$GarageFinish=='RFn',2,3)))
train_data$GarageQual = ifelse(is.na(train_data$GarageQual),6,ifelse(train_data$GarageQual=='Ex',1,
                      ifelse(train_data$GarageQual=='Gd',2,
                      ifelse(train_data$GarageQual=='TA',3,
                      ifelse(train_data$GarageQual=='Fa',4,5)))))
train_data$GarageCond = ifelse(is.na(train_data$GarageCond),6,
                      ifelse(train_data$GarageCond=='Ex',1,
                      ifelse(train_data$GarageCond=='Gd',2,
                      ifelse(train_data$GarageCond=='TA',3,
                      ifelse(train_data$GarageCond=='Fa',4,5)))))
train_data$PavedDrive = ifelse(train_data$PavedDrive=='Y',1,
                      ifelse(train_data$PavedDrive=='P',2,3))
train_data$PoolQC = ifelse(is.na(train_data$PoolQC),5,ifelse(train_data$PoolQC=='Ex',1,
                      ifelse(train_data$PoolQC=='Gd',2,
                      ifelse(train_data$PoolQC=='TA',3,4))))
train_data$Fence = ifelse(is.na(train_data$Fence),5,
                      ifelse(train_data$Fence=='GdPrv',1,
                      ifelse(train_data$Fence=='MnPrv',2,
                      ifelse(train_data$Fence=='GdWo',3,4))))
```
check structure of dataframe after change to nominal
```{r}
a = colSums(is.na.data.frame(train_data))
a[a>0]
```
make sure number of numeric columns is higher
```{r}
str(train_data)
```

now recalculate the correlation matrix
```{r}
library(corrplot)
library (corrr)
train_data_num = train_data %>% select_if(., is.numeric)
numcor = cor(train_data_num)
write.csv(numcor,'numcor.csv')
```

```{r}
i = 1
j = 1
for (i in 1:dim(numcor)[1]){
  for (j in 1:dim(numcor)[1]){
    if (is.na(numcor[i,j])){}
    else if(abs(numcor[i,j])>0.7 & abs(numcor[i,j])<0.999) {print(paste(
      rownames(numcor)[i],
      colnames(numcor)[j],
      numcor[i,j]
    ))}
  }
}
```
see if we can find correlations between categoricals and numerics with anova

```{r}
train_data2 = train_data %>% select(., -c(Alley, MiscFeature))
train_data2 = train_data2 %>% select(., -c(LotFrontage, BsmtFinType2,GarageType, GarageYrBlt, SalePrice))
colSums(is.na(train_data2))
train_numrc = train_data2 %>% select_if(., is.numeric)
train_chrc = train_data2 %>% select_if(., is.character)
# str(train_chrc)
# str(train_numrc)
```



```{r}
# for (i in 1:dim(train_chrc)[2]) {
#   print(colnames(train_chrc)[i])
#   print(table(train_chrc[,i]))
# }
omts = 0
b=''
for (i in 1:dim(train_chrc)[2]){
  if(sum(is.na(train_chrc[,i]))>omts | min(table(train_chrc[,i]))<2){
    b = c(b,(paste(i,j,'fail for na in category or too few in one category')))
  }
  else {
      for (j in 2:dim(train_numrc)[2]){
        if(sum(is.na(train_numrc[i,j]))>omts){
          b = c(b,(paste(i,j,'fail for na in numeric')))
        }
        else{
          a = oneway.test(train_numrc[ ,j]~train_chrc[ ,i], na.action = na.omit)
          if (is.na(a$p.value)) {}
          else if (a$p.value<0.01) {
            b = c(b,(paste(i,j,a$p.value,colnames(train_numrc)[j],colnames(train_chrc)[i])))
        }
        }
      }
  }
}
```
write p values to file
```{r}
write.csv(b,'anovas.csv')
```

pull out columns that correlated to other columns
```{r}
train_data3 = train_data %>% select(., -c('Id','ExterQual','TotalBsmtSF','TotRmsAbvGrd', 'FireplaceQu','GarageArea','GarageCond','PoolArea','MSZoning','BldgType','HouseStyle','CentralAir','MiscFeature','Alley','LotFrontage','GarageType','GarageYrBlt'))
train_data3
```
fit a model
```{r}
# colSums(is.na(train_data3))[colSums(is.na(train_data3))>0]
train_data3 = train_data3[complete.cases(train_data3),]
colSums(is.na(train_data3))[colSums(is.na(train_data3))>0]
```

```{r}
housemodel = glm(SalePrice~.,data = train_data3)
summary(housemodel)
plot(housemodel)
```

```{r}
housemodel
```

```{r}
housemodel$coefficients[housemodel$coefficients>500000]
plot(housemodel$coefficients)
```


```{r}
library(car)
vif(housemodel)
```









